A Formal Model of Checked C

### 2023-01-16 sandp23/implementation.tex
edit 5.3
Fixing eval
Shortening Implenentation section

\subsection{\sourcerewriter} \label{subsec:checkmate}
The~\sourcerewriter is primarily implemented in C++ as a~\clang frontend tool (\numcheckmatelines{}).
However, we use a small OCAML program (\numocamllines{}) to remove annotations and make the code compilable with the sandbox compiler (\sect{subsubsec:genucregion}).
\aravind{Arun, please fix the number of lines you changed for clang frontend and ocaml program}

@@

\noindent\emph{\_Mirror}: 
This qualifier makes ~\sourcerewriter copy the corresponding function into both ~\cregion and~\ucregion code such that callers from each region can invoke the corresponding copy.
This helps handle certain simple utility functions that might be called from both ~\cregion and~\ucregion code. For example, ~\code{json_value_get_type} in our evaluation of parsons\_wasm has callers from both the ~\cregion and ~\ucregion. 
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
/* JSON Value API */
_Mirror JSON_Value_Type json_value_get_type  
      (_TPtr<const TJSON_Value> value) {
    return value ? value->type : JSONError;
}

_Mirror _TPtr<TJSON_Object> json_value_get_object
(_TPtr<const TJSON_Value> value) {
    return json_value_get_type(value) == JSONObject ?
    value->value.object : NULL;
}

\end{minted}

Placing such common functions in either of the regions restricts their usage from the other region. For instance, if we place~\code{foo} in~\ucregion, then all parameters of it should be tainted. Consequently, callers from~\cregion should use tainted pointers to invoke~\code{foo}, making the arguments and other dependent pointers tainted, leading to over-tainting.
Similarly, placing~\code{foo} in~\cregion will make it hard for the callers in~\ucregion as they should now use the callback mechanism to invoke~\code{foo}.

Developers can use the ~\code{_Mirror} qualifier, which makes ~\sourcerewriter copy the function into both regions so that callers in each region will invoke the corresponding copy, thereby, avoiding unnecessary tainting.

However, ~\code{_Mirror} qualifier semantically asserts the function to only have callees that are either qualified with ~\code{_Tainted} or ~\code {_Mirror}. For example, ~\code{json_value_get_object}'s only callee 
\newline
~\code{json_value_get_type} is ~\code{\_Mirror} qualified.
Similarly, \systemname enforces ~\code{_Mirror} qualified functions to only access global checked data that are marked constant. These control-flow and data-flow compile-time semantics are enforced by \systemname to assert region locality and exclusivity amongst qualified function scopes in both the regions.

\noindent\emph{\_TLIB}:
This qualifier relaxes type-checking rules on library functions, thereby, avoiding unnecessary copying of tainted data.
Specifically, this annotation enables developers to pass tainted pointers as arguments to library functions in~\cregion.

For instance, consider that in~\cregion, we want to string concat~\code{n} characters from tainted pointer~\code{t_str} to~\code{dst} (a checked pointer).
We cannot use~\code{strncat} because our type-checker restricts this as it expects both the pointers to be non-tainted as we are in~\cregion.
We need to either implement the logic for string concatenation in~\cregion or first copy the contents of~\code{t_str} into a~\cregion temporary buffer (sat~\code{tmp}) and then invoke~\code{strncat(dst, tmp, n);}.
Both of these approaches are inefficient and, if not done correctly, might result in security bugs.
If the developer knows that~\code{strncat} accesses only~\code{n} bytes from the source buffer, she would like to directly use~\code{strncat} with tainted source pointer in~\cregion.
Consequently, ~\code{\_TLIB} qualified functions originally implemented in ~\cregion are expected to be non memory-modifying and thereby, its safe use is left at developer's discretion as \systemname does not enforce any semantics for this.  For example, malloc cannot be qualified with ~\code{_TLIB} and thereby is implemented as an indirect call to sandbox-specific allocator.

The~\code{_TLIB} qualifier enables the qualified function to be used as shown below:
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}

_TLIB extern char *strcat (char *__restrict __dest,
            const char *__restrict __src); // In the header file
...
// First, manually check the memory is in tainted region.
// if yes, then call strncpy.
if (!is_mem_in_range(t_str, t_str + n, SBX_LOW(), SBX_HIGH()))
  handle_violation();
// our type checker ignores this because 
// the above _TLIB annotation.
strncat(dst, t_str, n);
\end{minted}
The above qualifier relaxes type-checking for all the arguments to the calls to~\code{strncat}, enabling the developer to use the function freely in~\cregion.
Note that developer needs to manually add the check before calling~\code{strcat}.

@@

\subsection{Having Non-sandboxed Pointers}
\input{figures/newusecase}

>>>>

\subsection{\sourcerewriter} \label{subsec:checkmate}
The~\sourcerewriter is primarily implemented in C++ as a~\clang frontend tool (\numcheckmatelines{3028}).
However, we use a small OCAML program (\numocamllines{680}) to remove annotations and make the code compilable with the sandbox compiler (\sect{subsubsec:genucregion}).
\aravind{Arun, please fix the number of lines you changed for clang frontend and ocaml program}

@@

\noindent\emph{\_Mirror}: 
This qualifier makes ~\sourcerewriter copy the corresponding function into both ~\cregion and~\ucregion code such that callers from each region can invoke the corresponding copy.
This helps handle certain simple utility functions that might be called from both regions. For example, ~\code{append_string} in our evaluation of parsons\_wasm has callers from both the regions. 
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
_Mirror int append_string(_TPtr<char> buf,
const char* appendStr : itype(_Nt_array_ptr<const char>),
_TPtr<char> buf_start, size_t buf_len) {
/* Qualifier Rules:
1.) No access to global data NOT marked "const"
2.) Callees must be _Tainted or _Mirror
*/
...
}

\end{minted}
Placing ~\code{append_string} in ~\ucregion would require marking ~\code{appendStr} as tainted, thereby, requiring callers from ~\cregion to pass tainted parameter to ~\code{appendStr} at call-site. This would require either costly marshalling or making the arguments and other dependent pointers tainted, leading to over-tainting.
Similarly, placing ~\code{append_string} only in ~\cregion will make it hard for the callers in ~\ucregion as they should now use the callback mechanism to invoke~\code{append_string}.
Qualifying ~\code{append_string} with~\code{_Mirror} makes ~\sourcerewriter duplicate the function to ~\ucregion so that callers in each region will invoke their corresponding local copy, thereby, avoiding unnecessary tainting. 
"\_Mirror" enforces control-flow and data-flow compile-time semantic rules as shown above to assert region locality and exclusivity amongst between both the regions.

\noindent\emph{\_TLIB}:
This qualifier relaxes type-checking rules on non-memory modifying library functions allowing developers to use the function freely in ~\cregion.
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
// First, manually check the memory is in tainted region.
// if yes, then call strncpy.
if (!is_mem_in_range(t_str, t_str + n, SBX_LOW(), SBX_HIGH()))
  handle_violation();
// our type checker ignores this because 
// the above _TLIB annotation.
strncat(dst, t_str, n);
\end{minted}

In the above code, passing tainted pointer ~\code{t_str} to unqualified ~\code{strncat} is disallowed by our type-checker as it expects both the pointers to be non-tainted as we are in ~\cregion. If the developer knows that~\code{strncat} accesses only~\code{n} bytes from the source buffer, she would like to directly use ~\code{strncat} with tainted source pointer in ~\cregion, instead of implementing the logic for string concatenation in ~\ucregion or marshalling data between the two regions.  
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
- extern char *strncat (char *__restrict __dest,
+ _TLIB extern char *strncat (char *__restrict __dest,
const char *__restrict __src, size_t __n); // In the header file
}
\end{minted}

Consequently, annotating ~\code{strncat} as shown above relaxes type-checking for all the arguments to its calls. However, \systemname does not enforce any semantics to ensure ~\code{\_TLIB} functions implemented in ~\cregion are non memory-modifying, thereby, requiring developers to be careful to while using ~\code{\_TLIB} to prevent memory address leaks.

@@

\subsection{Balance Assurance and Overhead}
\input{figures/newusecase}
